---
layout: page
title: Exercise - Data Structures
author: Derek H. Ogle
css: "/css/modules.css"
output:
  html_document:
    self_contained: false
    lib_dir: zlibs
---

```{r echo=FALSE, eval=FALSE}
# Renders an appropriate HTML file for the webpage (CTRL-S and CTRL-ALT-R)
source(file.path(here::here(),"rhelpers/rhelpers.R"))
fnm <- "CE_1"
dnm <- "modules/Data-Structures"
modHTML(file.path(here::here(),dnm,fnm))
FSA::purl2(file.path(here::here(),dnm,paste0(fnm,".Rmd")),
           moreItems=c("source","setwd","fnm","dnm","opts_chunk","CTRL-S"))
```
```{r echo=FALSE, results='hide', message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo=FALSE, results="hide")
library(tidyverse)
```

----

## Vectors
1. Create four vectors each with at least five values of your choosing. Each vector must have a different class type and should be assigned to an object with an (at least semi-) informative name. Show the code that created the vector, showed the contents of the vector, and demonstrated the class of the vector.

```{r}
species <- c("Bluegill","Pumpkinseed","Black Crappie","Yellow Perch","Rockbass")
species
class(species)
caught <- c(15L,3L,7L,32L,3L)
caught
class(caught)
cpe <- c(4.5,0.7,1.5,9.1,0.6)
cpe
class(cpe)
happy <- c(TRUE,TRUE,FALSE,TRUE,FALSE)
happy
class(happy)
```

2. With one of your vectors from the first question, demonstrate code that will show one specific item in the vector.

```{r}
species[2]
caught[3]
cpe[4]
happy[1]
```

3. With one of your vectors from the first question, demonstrate code that will show multiple specific items in the vector.

```{r}
species[c(3,4)]
caught[c(4,5,1)]
```

&nbsp;

## Data Frames I
1. Create a data frame that has at least three related variables recorded on at least five individuals. You may use the same variables/data as in the "Vectors" question if the data in your vectors are related. Show the code that created the data frame, showed the full contents of the data frame, and showed the structure of the data frame.

```{r}
fish <- data.frame(species,caught,cpe,happy)
fish
str(fish)
```

2. Show how to find one specific value in your data frame.

```{r}
fish[1,2]
fish[2,3]
```

3. Show how to isolate a single row from your data frame.

```{r}
fish[2,]
```

4. Show how to isolate multiple rows from your data frame.

```{r}
fish[c(2,3),]
```

5. Show three methods to isolate a single variable from your data frame.

```{r}
fish[,1]
fish[,"species"]
fish$species
```

&nbsp;

## Data Frames II
Enter the following credit load by year-in-school data into a data frame in tidy format. Show the code that created the data frame, showed the full contents of the data frame, and showed the structure of the data frame.

```
First-Year  12  14  17  12
Sophomore   15  16  14
Junior      13  16  15  15  13
Senior      12  17
```

```{r}
students <- data.frame(yr=c("First-Year","First-Year","First-Year","First-Year",
                            "Sophomore","Sophomore","Sophomore",
                            "Junior","Junior","Junior","Junior","Junior",
                            "Senior","Senior"),
                       creds=c(12,14,17,12,
                               15,16,14,
                               13,16,15,15,13,
                               12,17))
students
str(students)
```

## Data Frames III

