---
layout: page
title: Exercise - Reading External Files
author: Derek H. Ogle
css: "/css/modules.css"
output:
  html_document:
    self_contained: false
    lib_dir: zlibs
---

```{r echo=FALSE, eval=FALSE}
# Renders an appropriate HTML file for the webpage (CTRL-S and CTRL-ALT-R)
source(file.path(here::here(),"rhelpers/rhelpers.R"))
fnm <- "CE_1"
dnm <- "modules/Reading-External-Files"
modHTML(file.path(here::here(),dnm,fnm))
FSA::purl2(file.path(here::here(),dnm,paste0(fnm,".Rmd")),
           moreItems=c("source","setwd","fnm","dnm","opts_chunk","CTRL-S"))
```
```{r echo=FALSE, results='hide', message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo=FALSE, results="hide")
library(tidyverse)
```

----

## Enter Data I

1. Enter data of your choosing into a spreadsheet (preferably Excel) for at least three related variables recorded on at least five individuals. You may use the same data as in your [Data Structures Exercise](../Data-Structures/CE_1.html#data-frames-i). Save the file in your spreadsheet software and as a CSV file.

1. Read your external CSV file into a data frame object in R. Show your code for reading the external file, for showing the full data frame, and for showing the structure of the data frame.

1. Read your external Excel file into a data frame object in R. [*If you used Google Sheets instead of Excel then load directly from Google Sheets.*] Show your code for reading the external file, for showing the full data frame, and for showing the structure of the data frame.

&nbsp;

## Enter Data II
Enter the credit load by year-in-school data from [this Data Structures Exercise](../Data-Structures/CE_1.html#data-frames-ii) into a data frame in tidy format. Show your code for reading the external file, for showing the full data frame, and for showing the structure of the data frame.

&nbsp;

## Reading Data I
Read [these data](https://raw.githubusercontent.com/droglenc/NCWrangling/main/resources/data/ChallengeData_1.csv) into R. Show your code for reading the external file, for showing the full data frame, and for showing the structure of the data frame. [*You should examine the external file prior to trying to read it into R.*]

