---
layout: page
title: Exercise - Reading External Files
author: Derek H. Ogle
css: "/css/modules.css"
output:
  html_document:
    self_contained: false
    lib_dir: zlibs
---

```{r echo=FALSE, eval=FALSE}
# Renders an appropriate HTML file for the webpage (CTRL-S and CTRL-ALT-R)
source(file.path(here::here(),"rhelpers/rhelpers.R"))
fnm <- "CE_1"
dnm <- "modules/Reading-External-Files"
modHTML(file.path(here::here(),dnm,fnm))
FSA::purl2(file.path(here::here(),dnm,paste0(fnm,".Rmd")),
           moreItems=c("opts_chunk"))
```
```{r echo=FALSE, results='hide', message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo=FALSE,results="hide",message=FALSE)
library(tidyverse)
```

----

## Enter Data I

1. Enter data of your choosing into a spreadsheet (preferably Excel) for at least three related variables recorded on at least five individuals. You may use the same data as in your [Data Structures Exercise](../Data-Structures/CE_1.html#data-frames-i). Save the file in your spreadsheet software and as a CSV file.

2. Read your external **CSV** file into a data frame object in R. Show your code for reading the external file, for showing the full data frame, and for showing the structure of the data frame.

```{r}
df1c <- read_csv(file.path("..","..","resources","data","FishData.csv"))
df1c
str(df1c)
```

3. Read your external **Excel** file into a data frame object in R. [*If you used Google Sheets instead of Excel then load directly from Google Sheets.*] Show your code for reading the external file, for showing the full data frame, and for showing the structure of the data frame.

```{r}
df1e <- readxl::read_excel(file.path("..","..","resources","data","FishData.xlsx"))
df1e
str(df1e)
```

&nbsp;

## Enter Data II
Enter the credit load by year-in-school data from [this Data Structures Exercise](../Data-Structures/CE_1.html#data-frames-ii) into a data frame in tidy format. Show your code for reading the external file, for showing the full data frame, and for showing the structure of the data frame.

```{r}
df2 <- read_csv(file.path("..","..","resources","data","StudentData.csv"))
df2
str(df2)
```

&nbsp;

## Reading Data I
Read [these data](https://raw.githubusercontent.com/droglenc/NCWrangling/main/resources/data/ChallengeData_1.csv) into R. Show your code for reading the external file, for showing the full data frame, and for showing the structure of the data frame. [*You should examine [the external file](https://github.com/droglenc/NCWrangling/blob/main/resources/data/ChallengeData_1.csv) prior to trying to read it into R.*]

```{r}
cdf1 <- read_csv("https://raw.githubusercontent.com/droglenc/NCWrangling/main/resources/data/ChallengeData_1.csv",skip=4)
cdf1
str(cdf1)
```

&nbsp;

## Reading Data II
Read [these data](https://raw.githubusercontent.com/droglenc/NCWrangling/main/resources/data/ChallengeData_2.csv) into R. Show your code for reading the external file, for showing the full data frame, and for showing the structure of the data frame. [*You should examine [the external file](https://github.com/droglenc/NCWrangling/blob/main/resources/data/ChallengeData_2.csv) prior to trying to read it into R.*]

```{r}
cdf2 <- read_csv("https://raw.githubusercontent.com/droglenc/NCWrangling/main/resources/data/ChallengeData_2.csv")
cdf2
str(cdf2)
```

&nbsp;

## Reading Data III
Read [these data](https://raw.githubusercontent.com/droglenc/NCWrangling/main/resources/data/ChallengeData_3.csv) into R. Show your code for reading the external file, for showing the full data frame, and for showing the structure of the data frame. [*You should examine [the external file](https://github.com/droglenc/NCWrangling/blob/main/resources/data/ChallengeData_3.csv) prior to trying to read it into R.*]

```{r}
cdf3 <- read_csv("https://raw.githubusercontent.com/droglenc/NCWrangling/main/resources/data/ChallengeData_3.csv",comment="##")
cdf3
str(cdf3)
```

&nbsp;

## Reading Data IV
Read [these data](https://raw.githubusercontent.com/droglenc/NCWrangling/main/resources/data/ChallengeData_4.csv) into R by making sure that all missing values are treated properly. Show your code for reading the external file, for showing the full data frame, and for showing the structure of the data frame. [*You should examine [the external file](https://github.com/droglenc/NCWrangling/blob/main/resources/data/ChallengeData_4.csv) prior to trying to read it into R.*]

```{r}
cdf4 <- read_csv("https://raw.githubusercontent.com/droglenc/NCWrangling/main/resources/data/ChallengeData_4.csv",na="XXX")
cdf4
str(cdf4)
```

&nbsp;

## Reading Data V
Read [these data](https://raw.githubusercontent.com/droglenc/NCWrangling/main/resources/data/ChallengeData_5.csv) into R by making sure that all missing values are treated properly. Show your code for reading the external file, for showing the full data frame, and for showing the structure of the data frame. [*You should examine [the external file](https://github.com/droglenc/NCWrangling/blob/main/resources/data/ChallengeData_5.csv) prior to trying to read it into R.*]

```{r}
cdf5 <- read_csv("https://raw.githubusercontent.com/droglenc/NCWrangling/main/resources/data/ChallengeData_5.csv",na=c("NA","NONE"))
cdf5
str(cdf5)
```

&nbsp;

## Reading Data VI
Read [these data](https://raw.githubusercontent.com/droglenc/NCWrangling/main/resources/data/ChallengeData_6.csv) into R by making sure that all missing values are treated properly. Show your code for reading the external file, for showing the full data frame, and for showing the structure of the data frame. [*You should examine [the external file](https://github.com/droglenc/NCWrangling/blob/main/resources/data/ChallengeData_6.csv) prior to trying to read it into R.*]

```{r}
cdf6 <- read_csv("https://raw.githubusercontent.com/droglenc/NCWrangling/main/resources/data/ChallengeData_6.csv",na="NONE",skip=6)
cdf6
str(cdf6)
```

&nbsp;

## Reading Data VII
Read the sheet named "ChallengeData_6" in [these data](https://github.com/droglenc/NCWrangling/raw/main/resources/data/ChallengeData.xlsx) into R by making sure that all missing values are treated properly. Show your code for reading the external file, for showing the full data frame, and for showing the structure of the data frame. [*You should examine the external file prior to trying to read it into R.*]

```{r}

cdf7 <- readxl::read_excel(file.path("..","..","resources","data","ChallengeData.xlsx"),
                           sheet="ChallengeData_6",na="NONE",skip=6)
cdf7
str(cdf7)
```
